---
title: "Geospatial Computing — Course Concept"
author: "Nils Ratnaweera"
date: "2026-04-23"
lang: en
format:
  typst:
    toc: true
    number-sections: true
execute: 
  echo: false
---


## Module at a Glance

|                        |                                                    |
|------------------------|----------------------------------------------------|
| **Study programme**    | BSc Umweltingenieurwesen, ZHAW IUNR                |
| **Type**               | Wahlpflichtmodul (Minor: Spatial Data Science)     |
| **Semester**           | 5th semester (HS 2026)                             |
| **ECTS**               | 5 (= 150 h total workload)                         |
| **Contact hours**      | ~4 × 45 min/week × 14 weeks ≈ 56 h                 |
| **Language**           | English                                            |
| **Prerequisite**       | Geoinformatik und GIS (3rd semester)               |
| **Students**           | ~30, environmental engineering                     |
| **Module coordinator** | Nils Ratnaweera                                    |
| **Co-lecturers**       | murj, rahn                                         |

## Learning Outcomes

Students are able to:

- process and analyse vector and raster geospatial data in a script-based workflow
- combine script-based geospatial workflows with GIS desktop software (QGIS)
- independently carry out spatial analyses and transformations of geospatial datasets
- develop automated and reproducible geospatial workflows
- visualise and communicate spatial information in a task-appropriate manner
- use AI-assisted tools to support coding and geospatial analysis tasks
- apply learned methods to practical problems in environmental and spatial analysis

## Course Structure

The course runs over 14 weeks and is organised around three thematic blocks, each pairing a **method focus** with an **application domain**. Each block has a main two-week session and one advanced follow-up session in weeks 9–11. The second half closes with a semester project phase.

```{r}
#| output: false
library(readxl)
library(tidyr)
library(dplyr)
library(gt)

plan <- read_xlsx(
  "Geospatial-Computing-CurrFS27.xlsx",
  sheet = "Wochenprogramm",
  n_max = 14
) |>
  select(`W#`, Block, Lecturer, Subtitle, Theory, `Seminar / Group Work`, `Labs / Exercises`) |>
  fill(Block, Lecturer)  # forward-fill merged cells from xlsx
```

```{r}
#| label: tbl-schedule
#| column: page
#| tbl-cap: Weekly Schedule
plan |>
  mutate(
    .run   = cumsum(Block != lag(Block, default = "") |
                    Lecturer != lag(Lecturer, default = "")),
    .group = paste0(Block, " (", Lecturer, ")", strrep("\u200b", .run))
  ) |>
  group_by(.group) |>
  mutate(across(everything(), \(x)ifelse(is.na(x),"",x))) |> 
  gt(rowname_col = "W#") |>
  cols_hide(c("Block", "Lecturer", ".run")) |>
  cols_label(
    Subtitle               = "Topic",
    Theory                 = "Theory",
    `Seminar / Group Work` = "Seminar",
    `Labs / Exercises`     = "Labs"
  ) |>
  tab_options(table.font.size = px(12))
```




## Semester Project

Students carry out a semester-long project in which they apply the methods from all three blocks to a dataset of their choice. The project runs in parallel with the teaching blocks and culminates in a written report.

The oral exam tests understanding of the methods and results — not the project output itself. This guards against unreflective use of AI tools.

## Assessment

| Component                                         | Type     | Weight |
|---------------------------------------------------|----------|--------|
| Project proposal                                  | Predicate (pass/fail) | — |
| Written project report                            | Predicate (pass/fail) | — |
| Coaching session attendance                       | Predicate (pass/fail) | — |
| Oral end-of-module exam (~20 min)                 | Grade    | 100%   |

All predicates must be passed to sit the oral exam. The module grade is determined solely by the oral exam.

**Logistics:** 30 students × 20 min = 10 hours of examining. This requires 2–3 exam days and a co-examiner.

