---
title: "Geospatial Computing — Course Planning Meeting"
subtitle: "Preparatory Information for Co-Lecturers"
author: "Nils Ratnaweera"
date: today
execute:
  echo: false
  eval: false
---

## Purpose of This Meeting

This document provides background and a proposed framework for the new *Geospatial Computing* module. The goal of the meeting is to agree on the course structure and divide topics between us three lecturers (murj, rahn, laup). 

## Module at a Glance

|                     |                                                             |
| ------------------- | ----------------------------------------------------------- |
| **Study programme** | BSc Umweltingenieurwesen, ZHAW IUNR                         |
| **Type**            | Elective module (Wahlpflichtmodul)                          |
| **Minor**           | Spatial Data Science                                        |
| **Semester**        | 5th semester (HS)                                           |
| **ECTS**            | 5 (= 150 h total workload)                                  |
| **Contact hours**   | ~4 × 45 min/week × 14 weeks ≈ 56 h                          |
| **Language**        | English                                                     |
| **Prerequisite**    | Geoinformatik und GIS (3rd semester)                        |
| **Students**        | ~30, environmental engineering, some programming background |

## Capacity Constraints

The 14-week semester is not fully available for teaching:

- ~1 week: project proposal presentations
- ~1 week: coaching sessions (mandatory attendance)

This leaves roughly **12–13 weeks for teaching inputs**, shared between three lecturers — approximately **3–4 weeks per person**.

## Assessment Structure (confirmed)

Students complete three pass/fail predicates and one graded exam:

| Component                         | Type      | Description |
| --------------------------------- | --------- | ----------- |
| Project proposal                  | Predicate | pass/fail   |
| Written project report            | Predicate | pass/fail   |
| Coaching session attendance       | Predicate | pass/fail   |
| Oral end-of-module exam (~20 min) | Grade     | 100% weight |

The oral exam tests understanding of methods and results, not the project output itself. This is a deliberate design choice to guard against unreflective AI use. All predicates must be passed to sit the exam.

## Proposed Course Framework

The course is organised along two axes: **methods** and **domains**.

### Methods Axis {#sec-methods}

Each lecturer teaches one or more method blocks. Each block is anchored to a specific application domain (see below), giving students a concrete dataset and problem to work with. The following methods are currently on the list — part of the meeting agenda is to reduce / enhance and confirm this list:

1. **Spatial pattern analysis** — autocorrelation, hotspot analysis, interpolation
2. **Uncertainty quantification** — Monte Carlo simulation, error propagation, sensitivity analysis
3. **Image processing** — automated detection from camera trap imagery (e.g. MegaDetector)
4. **Multi-Criteria Analysis (MCA)** — spatial suitability modelling, decision support
5. **Network Analysis** — routing, connectivity, accessibility

> We likely have capacity for **3–4 of these**, not all five. Selection is a key agenda item.

### Domains Axis {#sec-domain}

Each method block is taught in the context of one application domain. Students then choose a domain for their own semester project and apply the methods they have learned. Current candidate domains:

1. **Wildlife Monitoring** — camera trap data, opportunistic species observations
2. **Vegetation Ecology** — spatial and temporal vegetation surveys
3. **Hydrology** — vectorised river networks (no remote sensing or drone data)
4. **Environmental Planning** — visitor management, quiet area mapping

### Toolsets {#sec-toolsets}

The following tools are introduced where relevant, rather than as standalone blocks:

- R / RStudio
- QGIS
- QGIS Model designer
- Building a QGIS Plugin — Python-based extension
- Databases (PostGIS / SQLite)
- QField
- Webmapping

## Rough Semester Skeleton

@tbl-plan Shows a rough semester plan, very similar to the legacy AGI Modul. Each block is about 3 or 4 weeks long. Each block is in a [single domain]{.underline} and covers [1 or 2 methods]{.underline} introducing [1 or 2 toolsets]{.underline}.

```{r}
#| output: false
#| eval: false
library(tinytable)
library(dplyr)
library(readr)
semester_plan <- read_csv("semester_plan.csv") |> 
  mutate(
    lecturer_block = case_when(
      !is.na(lecturer)~paste(lecturer, block),
      .default = topic
    )
    ) |> 
  select(-c(lecturer, block)) |> 
  mutate(across(everything(), \(x)ifelse(is.na(x), "", x)))
```

```{r}
#| label: tbl-plan

select(semester_plan, -lecturer_block) |> 
  tt() |> 
  group_tt(i = semester_plan$lecturer_block)
```
## Course Framework: Three Axes

As described in @tbl-plan each block consists of one domain, and covers 1-2 methods and introduceds 1-2 toolsets. So we we should end up with about 1 - 3 lines per connection, as shown in below.

```{r}
#| fig-width: 9
#| fig-height: 5
#| warning: false
#| label: fig-connection 
#| fig-cap: Each topic connects the three axis, as shown in a single example.
library(ggplot2)
library(tibble)

domains  <- c("Wildlife Monitoring", "Vegetation Ecology", "Hydrology", "Environmental Planning")
methods  <- c("Spatial Pattern Analysis", "Uncertainty Quantification", "Image Processing", "Multi-Criteria Analysis", "Network Analysis")
toolsets <- c("R / Python", "QGIS", "QGIS Model Designer", "QGIS Plugin", "Databases", "QField", "Webmapping")

nodes <- bind_rows(
  tibble(col = 1, label = domains,  axis = "Domains"),
  tibble(col = 2, label = methods,  axis = "Methods"),
  tibble(col = 3, label = toolsets, axis = "Toolsets")
) |>
  group_by(col) |>
  mutate(y = (n():1) - (n() + 1) / 2) |>
  ungroup()

edges <- tribble(
  ~from,              ~to,
  "Network Analysis", "Environmental Planning",
  "Network Analysis", "R / Python"
) |>
  left_join(nodes |> select(label, x = col, y), by = c("from" = "label")) |>
  left_join(nodes |> select(label, xend = col, yend = y), by = c("to" = "label"))

ggplot(nodes, aes(x = col, y = y, label = label)) +
  geom_segment(data = edges, aes(x = x, y = y, xend = xend, yend = yend),
               inherit.aes = FALSE, colour = "steelblue", linewidth = 0.8) +
  geom_label(fill = "white", label.size = 0.3) +
  scale_x_continuous(breaks = 1:3, labels = c("Domains", "Methods", "Toolsets"),
                     limits = c(0.5, 3.5), position = "top") +
  theme_void() +
  theme(axis.text.x = element_text(face = "bold", size = 11))
```

## Agenda for Our Meeting


1. **Confirm the three lists**: Are any domains / methods / toolsets missing or superfluous?
2. **Perfect Matches**: 
   - Which domains, methods and toolsets offer themselves? 
   - What projects have we worked on that we could build upon?
   - (What datasets are suitable)
3. **What combinations of domains, methods and toolsets** are you intersted in following in your block? (A "connection" in @fig-connection)

